Characterising representable positive relation algebras with Priestley duality

### 2023-04-03 duality/Text.tex
minor updates and todos

\section{Priestley spaces for positive relation algebras}

A PRA-space is a structure $(X, \tau, \leq, R, I, E)$ where $X = (X, \tau, \leq)$ is a Priestley space and $R \subseteq X^3$, $I \subseteq X^2$ and $E \subseteq X$ such that:
\begin{itemize}
\item For all $x, y, z, w \in X$ there exists $u \in X$ such that $R(x, y, u)$ and $R(u, z, w)$ iff there exists $v \in X$ such that $R(y,z,v)$ and $R(x, v, w)$,
\item If $A, B \subseteq X$ are upward closed, so is $R[A, B, \underline{\:\:\:}]$,
\item $I(A)$ is upward closed clopen whenever $A$ is upward closed clopen,
\item $I(x)$ is closed for each $x \in X$,
\item For all $x, y, z \in X$, $x \leq y$ and $I(x, z)$ imply $I(y, z)$,
\item For all $x, y \in X$ there exists $z \in X$ such that $x = y$ iff $I(z, y)$ and $I(x, z)$,
\item For all $x, y, z \in X$ there exists $u \in X$ such that $I(u, z)$ and $R(x, y, u)$ iff there exist $u, w \in X$ such that $R(v, w, z)$, $I(y,v)$ and $I(x, w)$.
\item For all $x,y,u, v \in X$, $R(u,v,y)$ and $I(x, u)$ implies $R(x,y,v)$,
\item $E$ is upward closed clopen such that for each clopen $A \subseteq X$ one has
\begin{center}
$R[A, E, \overline{\:\:\:}] = R[E, A, \overline{\:\:\:}] = A$
\end{center}
\end{itemize}

\begin{fact} Let $\mathcal{R}$ be a positive relation algebra, then

\begin{enumerate}
\item If $F \subseteq \mathcal{R}$ is a filter, then for each $X, Y \subseteq \mathcal{R}$ $X ; Y \subseteq F$ iff $\uparrow (X ; Y) \subseteq F$
\item If $F_1, F_2 \subseteq \mathcal{R}$ are filters, so is $\uparrow (F_1 ; F_2)$,
\item Let $F_1, F_2$ be filters and $F_3 \in \operatorname{Spec}(\mathcal{R})$ such that $\uparrow (F_1 ; F_2) \subseteq F_3$, then there are $F_1', F_2' \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \subseteq F_1'$, $F_2 \subseteq F_2'$ and $\uparrow (F_1' ; F_2') \subseteq F_3$.
\end{enumerate}

\end{fact}

@@

\begin{enumerate}
\item $\mathcal{R} \hookrightarrow (2^{\operatorname{Spec}(R)}, \cup, \cap, \emptyset, \operatorname{Spec}(R), \bullet, \iota, \epsilon)$,
\item $\mathcal{R} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{R}))$,
\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
\item The categories of positive relation algebras and PRA-spaces are dually equivalent.
\end{enumerate}
\end{theorem}

>>>>

\subsection{Priestley duality for positive relation algebras}

Given a positive relation algebra $\mathcal{R} = (R, \cdot, +, ;, \smile, {\bf 1}', 0, 1)$, let $A, B \subseteq \mathcal{R}$, define the usual pointwise product operation:
\begin{center}
$A ; B = \{ a ; b \: | \: a \in A, b \in B \}$
\end{center}
We modify this operation with the upward closure:
\begin{center}
$A \bullet B = \uparrow (A ; B)$
\end{center}

We use the $\bullet$ operation to analyse the following properties of filters in positive relation algebras:
\begin{fact} Let $\mathcal{R}$ be a positive relation algebra, then

\begin{enumerate}
\item If $F \subseteq \mathcal{R}$ is a filter, then for each $X, Y \subseteq \mathcal{R}$ $X ; Y \subseteq F$ iff $X \bullet Y \subseteq F$
\item If $F_1, F_2 \subseteq \mathcal{R}$ are filters, so is $F_1 \bullet F_2$,
\item Let $F_1, F_2$ be filters and $F_3 \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \bullet F_2 \subseteq F_3$, then there are $F_1', F_2' \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \subseteq F_1'$, $F_2 \subseteq F_2'$ and $F_1' \bullet F_2' \subseteq F_3$.
\end{enumerate}
\end{fact}

First of all, given a positive relation algebra $\mathcal{R} = (R, \cdot, +, ;, \smile, 0, 1, {\bf 1}')$, we define its canonical extension $\mathcal{R}^{+}$ on the subsets of the spectrum $\operatorname{Spec}(\mathcal{R})$ by piggybacking Priestley representation of bounded distibutive lattices.

TODO:

A PRA-space is a structure $(X, \tau, \leq, R, I, E)$ where $X = (X, \tau, \leq)$ is a Priestley space and $R \subseteq X^3$, $I \subseteq X^2$ and $E \subseteq X$ such that:
\begin{enumerate}
\item For all $x, y, z \in X$ such that $R(x, y, z)$:
\begin{itemize}
\item $x \geq x'$ for $x' \in X$ implies $R(x', y, z)$,
\item $y \leq y'$ for $y' \in X$ implies $R(x, y', z)$,
\item $z \leq z'$ for $z' \in X$ implies $R(x, y, z')$.
\end{itemize}
\item For all $x, y, z, w \in X$ there exists $u \in X$ such that $R(x, y, u)$ and $R(u, z, w)$ iff there exists $v \in X$ such that $R(y,z,v)$ and $R(x, v, w)$,
\item If $A, B \subseteq X$ are upward closed, so is $R[A, B, \underline{\:\:\:}]$,
\item $I(A)$ is upward closed clopen whenever $A$ is upward closed clopen,
\item $I(x)$ is closed for each $x \in X$,
\item For all $x, y, z \in X$, $x \leq y$ and $I(x, z)$ imply $I(y, z)$,
\item For all $x, y \in X$ there exists $z \in X$ such that $x = y$ iff $I(z, y)$ and $I(x, z)$,
\item For all $x, y, z \in X$ there exists $u \in X$ such that $I(u, z)$ and $R(x, y, u)$ iff there exist $u, w \in X$ such that $R(v, w, z)$, $I(y,v)$ and $I(x, w)$.
\item For all $x,y,u, v \in X$, $R(u,v,y)$ and $I(x, u)$ implies $R(x,y,v)$,
\item $E$ is upward closed clopen such that for each clopen $A \subseteq X$ one has
\begin{center}
$R[A, E, \overline{\:\:\:}] = R[E, A, \overline{\:\:\:}] = A$
\end{center}
\end{enumerate}

@@

\begin{enumerate}
\item $\mathcal{R} \hookrightarrow \mathcal{R}^{+} = (2^{\operatorname{Spec}(R)}, \cup, \cap, \emptyset, \operatorname{Spec}(R), \bullet, \iota, \epsilon)$,
\item $\mathcal{R} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{R}))$,
\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
\item The categories of positive relation algebras and PRA-spaces are dually equivalent.
\end{enumerate}

\begin{proof} $ $

\begin{enumerate}
\item We check only the cases of $\bullet$, $\iota$ and $\epsilon$.
\item Note that the dual space of a positive relation algebra $\mathcal{R}$ is a PRA-space.

TODO: check the claim above

\item The dual space of a PRA-space is a positive relation algebra.

TODO: check the claim above

\item
\end{enumerate}
\end{proof}

\end{theorem}
